
---
title: "Business Data Analysis Report"
subtitle: "CSV Import and Data Cleaning"
author: "Helena Hall--665374757"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive Summary


## Question 1: CSV Import and Basic Exploration

Create a sample dataset to simulate importing CSV data:

```{r}
# Create sample regional sales data
regional_sales <- data.frame(
  region = c("North", "South", "East", "West", "North", "South", "East", "West",
             "North", "South", "East", "West", "North", "South", "East", 
             "North", "South","North"),  
  quarter = c(rep(c("Q1", "Q2", "Q3", "Q4"), each = 4), "Q4", "Q4"),
  revenue = c(125000, 98000, 115000, 102000, 
              138000, 105000, 128000, 118000,
              142000, 89000, 135000, 125000,
              155000, 112000, 148000, 165000, 
              120000, 91250),
  customers = c(450, 380, 420, 395, 
                485, 415, 465, 445,
                520, 350, 485, 470,
                575, 425, 545, 590, 435,324)
)
```

### Use str() to examine the structure of your data.

```{r}
str(regional_sales)
```

How many observations and variables do you have?

18 observations and 4 variables

What data types are the variables?

There are 2 types of variables:

chr = character/text

num = numeric (double precision numbers)


### Use head() and summary() to explore the data content.

```{r}
head(regional_sales)
```

```{r}
summary(regional_sales)
```


What is the range of revenue values?

89000-165000

Which region appears in the first few rows?

The first 4 rows show regions North, South, East and West


### Use unique() to identify all unique regions and quarters in your dataset.

```{r}
unique(regional_sales)
```

How many different regions and quarters are represented?

There are 4 different regions (North, South, East and West) and 3 quarters represented.

### Calculate the total number of customers across all regions and quarters using sum().

```{r}
sum(regional_sales$customers)
```

What business insight does this total provide?

This number shows how many customers were served in total in this time period. this can be helpful in comparing to previous years or to calculate average revenue per customer.

## Question 2: Data Quality Assessment

### Check for missing values using sum(is.na(regional_sales))

```{r}
sum(is.na(regional_sales))
```

Are there any missing values in this dataset?

No

Why is this check important in real business scenarios?

A missing value can impact the accuracy of decision made within the business and also reveal problems such as one area not reporting or a messed up POS system.

### Use table() to create frequency tables for the region and quarter variables.

```{r}
table(regional_sales$region, regional_sales$quarter)
```

Are the data evenly distributed across regions and quarters?

All data is evenly distributed throughout region and quarter for Quarter 1, 2 &3. In quarter 4, the data is not evenly distributed.

What does this tell you about your business operations?

This means for the first 3 quarters, the businesses operations were steady but in quarter 4 something changed to create an uneven distributions such as an expansion, or even gaps in data.

### Identify any potential data quality issues by finding records where revenue is less than $90,000. Use logical operators to filter and display these records with regional_sales[regional_sales$revenue < 90000, ]

```{r}
regional_sales[regional_sales$revenue < 90000, ]
```

Which region and quarter show concerning performance?

The South region in the 3rd quarter show just under $90000 of revenue.

## Question 3: Using Logical Operators for Business Questions

### Find all records where revenue is greater than $140,000 using regional_sales[regional_sales$revenue > 140000, ]

```{r}
regional_sales[regional_sales$revenue > 140000, ]
```

Which regions and quarters achieved high performance?

The North region achieved the highest performance in quarters 3 and 4 and were closely followed by the East region in quarter 4.

Count how many high-performance records exist using nrow().

```{r}
nrow(regional_sales[regional_sales$revenue > 140000, ])
```

### Identify "Q4" records with revenue greater than $130,000 AND more than 500 customers using the & operator.

```{r}
regional_sales[regional_sales$quarter == "Q4" & 
               regional_sales$revenue > 130000 & 
               regional_sales$customers > 500, ]
```

Display these records and explain what this combination tells you about year-end performance.

Quarter 4 shows a high revenue and high customer base showing a strong performance especially in the North and East region.

### Find records from either the "North" OR "East" regions using the | operator. Calculate the average revenue for these combined regions using mean(). 

```{r}
mean(regional_sales$revenue[regional_sales$region == "North" | 
                            regional_sales$region == "East"])
```

How does this compare to the overall average revenue?

The average overall revenue was $122847. The average revenue of just the North and East regions is higher than the overall average revenue.

### Create a business performance flag by identifying records that have EITHER revenue below $100,000 OR fewer than 400 customers.

```{r}
underperform_flag <- regional_sales$revenue < 100000 | 
                     regional_sales$customers < 400
```

```{r}
regional_sales[regional_sales$performance_flag == TRUE, ]
```

Count these concerning records and calculate what percentage of your total records they represent.

```{r}
sum(regional_sales$performance_flag)
```

```{r}
sum(regional_sales$performance_flag) / nrow(regional_sales) * 100
```

## Question 4: Creating Business Summaries

### Use aggregate() to calculate total revenue by region: aggregate(revenue ~ region, data = regional_sales, sum).

```{r}
aggregate(revenue ~ region, data = regional_sales, sum)
```

Which region generated the highest total revenue across all quarters?

North ($816250)

### Calculate average customers by quarter using aggregate(customers ~ quarter, data = regional_sales, mean).

```{r}
aggregate(customers ~ quarter, data = regional_sales, mean)
```

Which quarter shows the highest average customer count?

Quarter 4 (482.333 customers)

What business factors might explain this pattern?

This constant rise in customer counts could mean the business is growing through marketing efforts, expansion and starting to gain a reputation and strong customer base.

## Question 5: Executive Summary Creation

### Create a comprehensive business summary by calculating: total revenue across all regions and quarters, average revenue per quarter, and the number of high-performance records (revenue > $140,000). Present these as a formatted table or clear summary.

```{r}
# Total revenue
total_revenue <- sum(regional_sales$revenue)

# Average revenue per quarter
avg_revenue_quarter <- aggregate(revenue ~ quarter, data = regional_sales, mean)

# Number of high-performance records (revenue > 140000)
n_high_perf <- sum(regional_sales$revenue > 140000)

# Create a summary table
business_summary <- data.frame(
  Metric = c("Total Revenue", 
             paste("Average Revenue", avg_revenue_quarter$quarter, sep=" - "), 
             "High-Performance Records (Revenue > $140,000)"),
  Value = c(total_revenue, round(avg_revenue_quarter$revenue, 2), n_high_perf)
)

# View summary table
business_summary
```

### Based on your analysis, write a brief business recommendation (3-4 sentences) identifying which region shows the most consistent performance and which quarter typically shows the strongest results. Support your recommendation with specific numbers from your analysis.

The North region shows the most consistent performance, with multiple high-revenue records including two Q4 entries above $155,000 and over 575 customers each. Quarter 4 consistently generates the strongest results, with the highest average customer counts and several high-performance revenue records. It is recommended to focus marketing and operational resources on the North region during Q4 to maximize revenue, and consider applying successful Q4 strategies to other regions.