---
title: "Test 3"
author: "Helena Hall"
date: "2025-09-25"
output: html_document
---

---
title: "Business Insights Report"
subtitle: "Data Visualization and Communication"
author: "Helena Hall--665374757"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
# Then load the library once (and only once) in your RMD file
library(ggplot2)

##Data Preparation and Initial Analysis
```

```{r}
# Load ggplot2
library(ggplot2)

# Load the business dataset
business_data <- read.csv("data/business_performance.csv")
```


# Data preparation and Initial Analysis

```{r}
# Check the structure of the data
str(business_data)

# Get a summary of the data
summary(business_data)
```

## What type of data are the variables (columns) quarter and and channel?

- Quarter: character (categorical data)

- Channel: character (categorical data)

- Range of revenue values: 150,231 to 412,125


```{r}
# Calculate total revenue by channel
total_revenue_by_channel <- aggregate(revenue ~ channel, data = business_data, FUN = sum)

# View the results
total_revenue_by_channel
```

```{r}
# Find the channel with the highest total revenue
total_revenue_by_channel[which.max(total_revenue_by_channel$revenue), ]
```


```{r}
# Calculate average customers by quarter
avg_customers_by_quarter <- aggregate(customers ~ quarter, data = business_data, FUN = mean)

# View the results
avg_customers_by_quarter
```

```{r}
# Find the quarter with the highest average customers
avg_customers_by_quarter[which.max(avg_customers_by_quarter$customers), ]
```

```{r}
# Create revenue per customer column
business_data$revenue_per_customer <- business_data$revenue / business_data$customers

# View the first few rows to check
head(business_data)
```

```{r}
# Calculate average revenue per customer by channel
avg_rev_per_customer_by_channel <- aggregate(revenue_per_customer ~ channel, 
                                              data = business_data, 
                                              FUN = mean)

# View the results
avg_rev_per_customer_by_channel
```

```{r}
# Channel with the highest average revenue per customer
avg_rev_per_customer_by_channel[which.max(avg_rev_per_customer_by_channel$revenue_per_customer), ]
```

Business insight:


The channel with the highest revenue per customer is generating more revenue from each customer.


This can indicate:


- Higher-value customers using that channel


- More effective sales or upselling strategies


- Opportunity to focus marketing or resources on that channel to maximize profits.


# Creating Professional Charts

```{r}
# Calculate totals first
channel_totals <- aggregate(revenue ~ channel, data = business_data, sum)
# Then create the chart
ggplot(channel_totals, aes(x = channel, y = revenue)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Revenue by Channel", 
       x = "Sales Channel", y = "Revenue ($)")
```


## Which channel should receive more investment based on this visualization?


The Online channel should receive more investment because it has the highest total revenue

```{r}
ggplot(business_data, aes(x = customers, y = revenue, color = channel)) +
  geom_point(size = 3) +
  labs(title = "Customer Count vs Revenue by Channel",
       x = "Number of Customers", y = "Revenue ($)")
```

## Which channel appears most efficient at converting customers to revenue?


The online channel has this highest points overall which means it's generating more revenue per customer meaning its more efficient at converting cutomers into revenue



# Action Plan

```{r}
# Total revenue by channel
total_revenue <- aggregate(revenue ~ channel, data = business_data, sum)

# Average customers by channel
avg_customers <- aggregate(customers ~ channel, data = business_data, mean)

# Average conversion rate by channel
avg_conversion <- aggregate(conversion_rate ~ channel, data = business_data, mean, na.rm = TRUE)

# Combine into one table
summary_table <- merge(total_revenue, avg_customers, by = "channel")
summary_table <- merge(summary_table, avg_conversion, by = "channel")

# Optional: round numbers for readability
summary_table$customers <- round(summary_table$customers, 0)
summary_table$conversion_rate <- round(summary_table$conversion_rate, 1)

# View the table
summary_table
```



Invest more in the Online sales channel, which generates the highest total revenue of $5,571,130 and serves an average of 1,560 customers with a solid conversion rate of 3.6%. Focused marketing and resources here can maximize returns. Additionally, consider leveraging Retail strategies to improve conversion efficiency, as it has the highest conversion rate of 5.0% but fewer customers, to boost overall revenue without major new investment.


