
---
title: "Introduction to E-Business Analytics"
subtitle: "Performance Metrics Report"
author: "Helena Hall--665374757"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

# Executive Summary
The Management 230 class with Johan Latulippe covers topics of website creation and making professional documents by using the program "R" and "RStudio". So far we have gone over Quiz 1 and learned the basic skills of creating a document using "Rmarkdown". We have written some code and used it to "knit" a professional document to use to analyse data, create charts, and nterpret statistics. These are all things we can use to analys data for a manager in our workplace.

First, create your dataset using this code:

```{r}
max(# Create website performance dataset with business terminology
webperfo <- data.frame(
  engagement = airquality$Temp,
  performance = airquality$Ozone,
  server_load = airquality$Wind
), na.rm = TRUE)
```

## Question 1

### How many columns of data do you see, and what are their names? 
There are 3 columns of data and the are named "engagement, performance, server_load"
Why is it useful to see just the first few rows instead of printing all the data at once?
You can see the pattern or if there are any mistakes easily.
```{r}
head(webperfo)
```

### Which variable has missing values (NA's)?
Performance
What does the summary tell you about the range of your engagement scores (minimum to maximum)?
Min = 56.00, Max = 97.00
```{r}
summary(webperfo)
```
### Why might this number be important for a business manager planning website improvements?
153 days of data could be useful information to a manager because it can let them know how reliable and representative the data is.
```{r}
nrow(webperfo)
```

### How does this compare to the median engagement shown in your summary?
The max is NA and the median engagement is 79.00
```{r}
max(webperfo)
```

### Why did we need to add na.rm = TRUE here but not when finding the maximum engagement?
Because if there are missing values we need to tell the computer to remove the missing values so it doesn't reply with "NA"
```{r}
max(webperfo$performance, na.rm = TRUE)
```

## Question 2

### Looking at this plot, can you see a general pattern between engagement and performance?
Yes, they both affect each other and as performance goes up, so does engagement
What makes this plot difficult to understand for a business presentation?
There are not many numbers of the axis and not very clear titles. We do not know what this plot is representing

```{r}
plot(webperfo$engagement, webperfo$performance)
```

### How do the title and axis labels make this plot more useful for a business audience?
Multiple people can look at the data with less context and have it make sense. Also keeps things organized knowing which plots refer to what.

```{r}
plot(x = webperfo$engagement, y = webperfo$performance,
     main = "User Engagement vs Website Performance",
     xlab = "Daily User Engagement Score", 
     ylab = "Website Performance Score",)
```

### What visual pattern do you now see between engagement and performance? Describe it in business terms.
Both performance and engagement trend upwards and their relationship with each other becomes less constant as they increase in value.

```{r}
plot(x = webperfo$engagement, y = webperfo$performance,
     main = "User Engagement vs Website Performance",
     xlab = "Daily User Engagement Score", 
     ylab = "Website Performance Score", 
     pch = 16, 
     col = "purple")
```


### What does the red trend line tell you? Is it going up, down, or staying flat as engagement increases?
The red trend line is showing an upward trend in the scatter chart following the upward trend of the data.

```{r}
plot(x = webperfo$engagement, y = webperfo$performance,
     main = "User Engagement vs Website Performance",
     xlab = "Daily User Engagement Score", 
     ylab = "Website Performance Score", 
     pch = 16, 
     col = "purple")
abline(lm(performance ~ engagement, data = webperfo), 
       col = "orange", lwd = 2)
```

## Question 3

### What is the correlation coefficient? Is this a strong (above 0.7), moderate (0.3-0.7), or weak (below 0.3) relationship?
This is a moderate relationship. "0.69" correlation

```{r}
cor(webperfo$engagement, webperfo$performance, use = "complete.obs")
```

### Based on your analysis, when engagement is high, what typically happens to website performance? What should an e-business manager do with this information?
When engagement is high, website performance improves, which is shown by the positive correlation of 0.69. This means that using better content to encourage user-engagement,improves overall performance. An e-business manager should prioritize strategies that drive engagement beacuse it would improve the website performance as well.
